Idealized Wind Models of Cool Gas Outflows

### 2010-09-10 Winds/papers/Theory/First/Submit/ms.tex
blah
git-svn-id: https://svn.ucolick.org/xsvn/trunk/radtrans@5191 ac11df5e-a788-c23b-a8e1-c758c323924d

\author{
J. Xavier Prochaska\altaffilmark{1}, 
Others
%John M. O'Meara\altaffilmark{2}, 
%Gabor Worseck\altaffilmark{1} 
%\& Scott Burles\altaffilmark{3}
}
\altaffiltext{1}{Department of Astronomy and Astrophysics, UCO/Lick Observatory, University of California, 1156 High Street, Santa Cruz, CA 95064}
%\altaffiltext{2}{Department of Chemistry and Physics, Saint Michael's College.
%One Winooski Park, Colchester, VT 05439}

@@

Another factor that may reduce line-emission is a
non-isotropic (e.g.\ biconical) wind model. The flux is lower, for example, if one
eliminates the backside to the wind (e.g.\ the source itself could
shadow a significant fraction of the backside).  
For our fiducial model, many of the photons scattered into our
sightline arise nearly normal to the sightline such that a biconicial
wind has signficantly lower line emission
($\S$~\ref{sec:anisotropic}). 
For the anisotropic winds explored in this paper,
the line-emission scales 
roughly as $\Omega/4\pi$ where $\Omega$ is the angular extent of the
wind\footnote{There is also a first order dependence on the actual
  orientation of the wind, with maximal emission when the backside is
  fully viewed.}.  Because we require that the wind 
points toward us, it is difficult to reduce $\Omega$ much
below $2 \pi$ and, therefore, anisotropy reduces the
emission by a factor of order unity.  
Similarly, an anisotropicly emitting source would modify the predicted
line-emission.  For example, if the backside of the galaxy were
brighter/fainter then would one predict brighter/fainter emission
relative to the observed continuum.
To reduce the emission, the brightest regions
of the galaxy would need to be oriented toward Earth.  Although this
would not be a
generic orientation, most spectroscopic samples are magnitude-limited
and biased towards detecting galaxies when observed from their
brightest viewpoint.  In principle, this might imply a further
reduction of order unity.

@@

The most obvious measures provided by line-emission analysis are
estimates
of the size and morphology of the outflow \citep{rubin+10c}.  
Line emission is predicted to extend to radii where the Sobolev
optical depth exceeds a few tenths.  The principle
challenge is to achieve sufficient sensitivity to detect low
surface brightness emission. As Figure~\ref{fig:obs_sb}
demonstrates, the surface brightness at the inner wind radius of our
fiducial wind exceeds that at the outer radius by several orders of magnitude.
Nevertheless, an instrument that sampled the entire wind (e.g.\ a
large format IFU or narrow band imager) may detect the emission in
azimuthally averaged apertures. {\bf [JXP: Give estimate]
[JXP: Add FeII* to the SB figure]}
Of the transitions considered in this paper, \ion{Mg}{2} emission is
preferred for this analysis because 
(i) it has the largest equivalent width in absorption and
(ii) there are fewer emission channels per absorption line than for
\ion{Fe}{2}.  The \ion{Mg}{2} lines (especially \mgiib) frequently has the
highest peak and integrated fluxes (Figure~\ref{fig:obs_sb}).  
On the other hand, the \ion{Mg}{2} transitions are more susceptible to
the effects of dust extinction and for some galaxies \feiis\ emission
will be dominant.  [Last comment]

The kinematics of the line-emission also offer insight into physical
characteristics of the wind.
In $\S$~\ref{sec:dust}, we emphasized that dust extinction
preferentially suppresses photons scattered off the backside of the
wind (e.g.\ with $\mdv > 0 \mkms$) so that the centroid is shifted to
negative velocities.  Similarly, anisotropic models with reduced
emission off the backside yield emission lines that are centered
blueward of the galaxy's systemic redshift.  These effects are mos
prominent in the \feiis\ emission.  The centroids of the \ion{Mg}{2}
lines, meanwhile, may be modififed by absorption, e.g.\ a wind with
speeds exceeding 600\kms\ will have \mgiib\ scatter \mgiia\ photons.
Therefore, the line centroids are sensitive to the optical depth
profile of the wind.

We also find that the emission-line widths trace specific
characteristics of the wind.  For the \feiis\ emission, the widths are
most sensitive to the optical depth at large velocities from
systemic.  This is illustrated in Figure~\ref{fig:edelv} which plots
the 90\%\ velocity width $\Delta v_e$ of \feiic\ versus the velocity
where the wind optical depth is maximal ($v_\tau$).  It is evident
that $\Delta v_e$ increases with $v_\tau$; a large measured width for
\feiis\ emission requires a wind profile with large optical depths at
$\mdv < 200 \mkms$.

@@

\begin{figure}
\includegraphics[scale=0.6,angle=90]{../Figures/fig_obs_slit.ps}
\caption{
Emission equivalent width ($W_{\rm e}$) relative to the observed
absorption equivalent width ($W_{\rm a}$) for a series of transitions:
(solid -- \mgiia; dotted -- \mgiib; dashed -- \feiib).  The
\ewe/\ewabs\ ratio is plotted as a function of slit width relative to
twice the radius \rtt, defined to be where the Sobolev optical depth 
$\tau^S = 0.2$.  The black curves correspond to the fiducial wind
model ($\S$~\ref{sec:fiducial}; $\mrtt \approx 15$\,kpc), the red curves are for the
LBG-partial covering scenario ($\S$~\ref{sec:Covering}; $\mrtt \approx
8$\,kpc), and the
radiative-driven wind ($\S$~\ref{sec:radiative}; $\mrtt = XXX$) has blue curves.
For all of the wind models, the \ewe/\ewabs\ ratio rises very steeply
with slit width and then plateaus at $\approx \mrtt$.  
Therefore, a slit that exceeds $\approx 0.5 \mrtt$ will admit nearlly
all of the photons scattered to our sightilne.
}
\label{fig:obs_slit}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.6,angle=90]{../Figures/fig_obs_peaktau.ps}
\caption{
The observed peak optical depth $\tau_{\rm pk} \equiv -\ln [I_{\rm
  min}]$ for \mgiia\ profiles of the wind models studied in this paper
(Tables~\ref{tab:line_diag}, \ref{tab:plaw_diag}) with $I_{\rm min}$
the minimum normalized intensity of the absorption profile.  Cases
where \tpk\ exceeds 3 are presented as lower limits.  The \tpk\ values
are plotted against the velocity $v_\tau$ where $I(v) = I_{\rm min}$.  In all
of the models, the true peak optical depth $\mtpk^{\rm true} > 10$.
The much lower `observed' \tpk\ values occur because scattered \mgiia\
photons have filled-in the absorption profiles at velocities $\mdv
\gtrsim -400 \mkms$. The effects are greatest for wind
models where the optical depth peaks near $\mdv \sim 0 \mkms$ because
the majority of scattered photons have this realtive velocity.
Indeed, models with $v_\tau < -400 \mkms$ all show $\mtpk > 3$.
}
\label{fig:obs_peaktau}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.6,angle=90]{../Figures/fig_obs_sb.ps}
\caption{
Surface brightness profiles of the fiducial wind model for the \mgiia\
(solid) and \feiic\ (dotted) transitions.  Both profiles peak at small
radii and decrease by several orders of magnitude before reaching the
outer edge of the wind.  The profiles are sufficiently shallow,
however, that the azimuthally integrated flux declines by only a
factor of XXX from $r = 1$\,kpc to $r=10$\,kpc.  Therefore, a
sensitive IFU observation could map the emission (and, in principle,
the kinematics) from $r_{\rm inner}$ to $r_{\rm outer}$.
}
\label{fig:obs_sb}
\end{figure}

\begin{figure}
\includegraphics[scale=0.6,angle=90]{../Figures/fig_obs_edelv.ps}
\caption{
Width of \feiis.
}
\label{fig:obs_edelv}
\end{figure}

>>>>

\author{
J. Xavier Prochaska\altaffilmark{1}, 
Daniel Kasen\altaffilmark{1,2,3}, 
Kate Rubin\altaffilmark{1} 
}
\altaffiltext{1}{Department of Astronomy and Astrophysics, UCO/Lick Observatory, University of California, 1156 High Street, Santa Cruz, CA 95064}
\altaffiltext{2}{UCB?}
\altaffiltext{3}{Hubble Fellow}

@@

Another factor that may reduce line-emission is a
non-isotropic (e.g.\ biconical) wind model. The flux is lower, for example, if one
eliminates the backside to the wind (e.g.\ the source itself could
shadow a significant fraction of the backside).  
For our fiducial model, many of the photons scattered into our
sightline arise nearly normal to the sightline such that a biconicial
wind has significantly lower line emission
($\S$~\ref{sec:anisotropic}). 
For the anisotropic winds explored in this paper,
the line-emission scales 
roughly as $\Omega/4\pi$ where $\Omega$ is the angular extent of the
wind\footnote{There is also a first order dependence on the actual
  orientation of the wind, with maximal emission when the backside is
  fully viewed.}.  Because we require that the wind 
points toward us, it is difficult to reduce $\Omega$ much
below $2 \pi$ and, therefore, anisotropy reduces the
emission by a factor of order unity.  
Similarly, an anisotropically emitting source would modify the predicted
line-emission.  For example, if the backside of the galaxy were
brighter/fainter then would one predict brighter/fainter emission
relative to the observed continuum.
To reduce the emission, the brightest regions
of the galaxy would need to be oriented toward Earth.  Although this
would not be a
generic orientation, most spectroscopic samples are magnitude-limited
and biased towards detecting galaxies when observed from their
brightest viewpoint.  In principle, this might imply a further
reduction of order unity.

@@

The most obvious measures provided by line-emission analysis are
estimates
of the size and morphology of the outflow \citep{rubin+10c}.  
Line emission is predicted to extend to radii where the Sobolev
optical depth exceeds a few tenths.  The principle
challenge is to achieve sufficient sensitivity to detect low
surface brightness emission. As Figure~\ref{fig:obs_sb}
demonstrates, the surface brightness at the inner wind radius of our
fiducial wind exceeds that at the outer radius by several orders of magnitude.
Nevertheless, an instrument that sampled the entire wind (e.g.\ a
large format IFU or narrow band imager) may detect the emission in
azimuthally averaged apertures. {\bf [JXP: Give estimate]}
Of the transitions considered in this paper, \ion{Mg}{2} emission is
preferred for this analysis because 
(i) it has the largest equivalent width in absorption and
(ii) there are fewer emission channels per absorption line than for
\ion{Fe}{2}.  The \ion{Mg}{2} lines (especially \mgiib) frequently has the
highest peak and integrated fluxes (Figure~\ref{fig:obs_sb}).  
On the other hand, the \ion{Mg}{2} transitions are more susceptible to
the effects of dust extinction and for some galaxies \feiis\ emission
will be dominant.  [Last comment]

The kinematics of the line-emission also offer insight into physical
characteristics of the wind.
In $\S$~\ref{sec:dust}, we emphasized that dust extinction
preferentially suppresses photons scattered off the backside of the
wind (e.g.\ with $\mdv > 0 \mkms$) so that the centroid is shifted to
negative velocities.  Similarly, anisotropic models with reduced
emission off the backside yield emission lines that are centered
blueward of the galaxy's systemic redshift.  These effects are mos
prominent in the \feiis\ emission.  The centroids of the \ion{Mg}{2}
lines, meanwhile, may be modified by absorption, e.g.\ a wind with
speeds exceeding 600\kms\ will have \mgiib\ scatter \mgiia\ photons.
Therefore, the line centroids are sensitive to the optical depth
profile of the wind.

We also find that the emission-line widths trace specific
characteristics of the wind.  For the \feiis\ emission, the widths are
most sensitive to the optical depth at large velocities from
systemic.  This is illustrated in Figure~\ref{fig:obs_edelv} which plots
the 90\%\ velocity width $\Delta v_e$ of \feiic\ versus the velocity
where the wind optical depth is maximal ($v_\tau$).  It is evident
that $\Delta v_e$ increases with $v_\tau$; a large measured width for
\feiis\ emission requires a wind profile with large optical depths at
$\mdv < 200 \mkms$.

@@

\begin{figure}
\includegraphics[scale=0.6,angle=90]{../Figures/fig_obs_slit.ps}
\caption{
Emission equivalent width ($W_{\rm e}$) relative to the observed
absorption equivalent width ($W_{\rm a}$) for a series of transitions:
(solid -- \mgiia; dotted -- \mgiib; dashed -- \feiib).  The
\ewe/\ewabs\ ratio is plotted as a function of slit width relative to
twice the radius \rtt, defined to be where the Sobolev optical depth 
$\tau^S = 0.2$.  The black curves correspond to the fiducial wind
model ($\S$~\ref{sec:fiducial}; $\mrtt \approx 15$\,kpc), the red curves are for the
LBG-partial covering scenario ($\S$~\ref{sec:Covering}; $\mrtt \approx
8$\,kpc), and the
radiative-driven wind ($\S$~\ref{sec:radiative}; $\mrtt = XXX$) has blue curves.
For all of the wind models, the \ewe/\ewabs\ ratio rises very steeply
with slit width and then plateaus at $\approx \mrtt$.  
Therefore, a slit that exceeds $\approx 0.5 \mrtt$ will admit nearly
all of the photons scattered to our sightline.
}
\label{fig:obs_slit}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.6,angle=90]{../Figures/fig_obs_peaktau.ps}
\caption{
The observed peak optical depth $\tau_{\rm pk} \equiv -\ln [I_{\rm
  min}]$ for \mgiia\ profiles of the wind models studied in this paper
(Tables~\ref{tab:line_diag}, \ref{tab:plaw_diag}) with $I_{\rm min}$
the minimum normalized intensity of the absorption profile.  Cases
where \tpk\ exceeds 3 are presented as lower limits.  The \tpk\ values
are plotted against the velocity $v_\tau$ where $I(v) = I_{\rm min}$.  In all
of the models, the true peak optical depth $\mtpk^{\rm true} > 10$.
The much lower `observed' \tpk\ values occur because scattered \mgiia\
photons have filled-in the absorption profiles at velocities $\mdv
\gtrsim -400 \mkms$. The effects are greatest for wind
models where the optical depth peaks near $\mdv \sim 0 \mkms$ because
the majority of scattered photons have this relative velocity.
Indeed, models with $v_\tau < -400 \mkms$ all show $\mtpk > 3$.
}
\label{fig:obs_peaktau}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.6,angle=90]{../Figures/fig_obs_sb.ps}
\caption{
Surface brightness profiles of the fiducial wind model for the \mgiia\
(solid) and \feiid\ (dotted) transitions.  Both profiles peak at small
radii and decrease by several orders of magnitude before reaching the
outer edge of the wind.  The profiles are sufficiently shallow,
however, that the azimuthally integrated flux declines by only a
factor of $\approx 10$ from $r = 1$\,kpc to $r=10$\,kpc.  Therefore, a
sensitive IFU observation could map the emission (and, in principle,
the kinematics) from $r_{\rm inner}$ to $r_{\rm outer}$.
}
\label{fig:obs_sb}
\end{figure}

\begin{figure}
\includegraphics[scale=0.6,angle=90]{../Figures/fig_obs_edelv.ps}
\caption{
Plot of the 90\%\ velocity width ($\Delta v_e$) of \feiic\ emission versus the
velocity where the resonance lines achieve peak optical depth
$v_\tau$ in absorption.  The $\Delta v_e$ values rise steadily with
increasing offset of $v_\tau$ from systemic.  The width of the \feiic\
emission line, therefore, offers an independent diagnostic of the wind
speed.
}
\label{fig:obs_edelv}
\end{figure}
